The Talking Heads Experiment

### 2015-04-10 chap7/chap7.tex
cumulative changes to preface and chapter 1
updated labels, image widths, quotation marks, capitalisation, etc. for chapters 1-7

%\documentclass{../_combined/fcg-book}
\chapter{Grounding}

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.50\textwidth]{chap7/figs/sieve3}}
\caption{ \label{sieve3} Flow of solutions through 
the different cognitive layers when an utterance is being produced.
Re-entry is necessary because the decision at each layer depends on
decisions at subsequent layers.} 
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[width=0.8\columnwidth]{chap7/figs/plate-10}
\end{center}
\caption{ \footnotesize Three examples of segmented images. The 
topic is indicated by a dashed bounding box in the 
image of the speaker. Segments which are too small 
are ignored. The topics have all been conceptualised
as being `to the right' and so the same word 
`gofubo' has been used to refer to them. }
\label{fig:plate-10}
\end{figure}

@@

\begin{table}
\begin{center}
\begin{tabular}{ l  l  l }
\lsptoprule
{\itshape channel}& {\itshape obj-0} & {\itshape obj-1}\\ \midrule
HPOS & 0.27 & 0.16\\ 
VPOS & 0.20 & 0.20\\ 
HEIGHT & 0.15 & 0.15\\ 
WIDTH & 0.10 & 0.11\\ 
AREA & 0.10 & 0.10\\ 
R & 0.23 & 0.25\\ 
G & 0.32 & 0.34\\ 
B & 0.63 & 0.65\\ 
\lspbottomrule
\end{tabular}
\caption{ \label{tab:game3b} First words after about a dozen games.}
\end{center}
\end{table}

HPOS is the most salient channel. After sensor-scaling, 
the two values for HPOS are still be drawn further
apart with 1.0 for object-0 and 
0.0 for object-1 so that the category [HPOS 0.5-1.0] easily 
distinguishes the topic (object-0) from object-1.

The left image is that of the hearer {\bfshape a1}, the right one
that of the speaker, {\bfshape a2}.
{\bfshape a2} had already invented the word `gofubo' for 
[HPOS 0.5-1.0] (to the right) in an earlier game, 
but the word was not acquired in that game
by {\bfshape a1} because he still missed the 
appropriate distinction. Meanwhile the HPOS category 
is available to {\bfshape a1} due to an
expansion of his discrimination networks
and so he can store the word: 
\begin{verbatim}
Game 3 
  a2 is the speaker. a1 is the hearer. 
  a2 segments the context into 2 objects: 
       object-0 object-1
  a2 chooses object-0 as the topic 
  a2 categorises the topic as [HPOS 0.5-1.0]
  a2 says: gofubo
  a1 does not know gofubo
  a1 says: gofubo?
  a2 points to object-0
  a1 categorises the topic as [HPOS 0.5-1.0]
  a1 stores gofubo as [HPOS 0.5-1.0]
\end{verbatim}
This game proceeds essentially like in the computer
simulations studied before. The major difference is
that now real images have been used, the objects
considered are the outcome of segmentation processes,
the sensory characteristics have been derived from the 
image itself, and the pointing has been done by 
physically moving the cameras.

@@

Immediately the agents apply this word to very 
different scenes, such as the ones in \figref{fig:plate-10} (middle and bottom). The middle picture
shows on the left the segmented scene
from the speaker in game 5, and on the right the one
from the hearer. In game 5, the figures are blue
rectangles. They are much further apart than 
the two circles in game 3. Nevertheless, 
after scaling, they are categorised and
conceptualised similarly and therefore
the same word could be used effectively. The concept 
of [HPOS 0.5-1.0] and hence the word `gofubo' 
is general {\itshape from the very start}. 
The agents do not need to see many examples because they do 
not use inductive generalisation. Instead, 
the HPOS category is constructed in a top-down fashion 
as soon as the HPOS channel has been salient
and is immediately available for use in the 
discrimination game and hence in verbalising 
the scene. This explains why the word-meaning acquisition
process observed in the experiments goes so 
amazingly fast.

The bottom of \figref{fig:plate-10}
shows yet another scene where the word `gofubo' 
was used with success. The topic is the rightmost
shape in the scene and so [HPOS 0.5-1.0] is once more 
distinctive. To an outside observer it may look 
like the agents have performed a gigantic inductive
leap, but this is not the case at all. The agents
do not try to abstract the commonalities from 
different examples but construct distinctions in a top 
down fashion and try to apply them to the perceived image.

@@

\begin{table}
\begin{center}
\begin{tabular}{ l  l  l  l  l }
\lsptoprule
{\itshape Word}&{\itshape Meaning}&{\itshape Translation} & {\bfshape a1}&{\bfshape a2} \\ \midrule
wawosido & [HPOS 0.0-0.5] &left&0.4&0.4\\ 
meluri & [HPOS 0.25-0.5] &medium left&0.1&0.0\\ 
gofubo & [HPOS 0.5-1.0]& right&1.0&1.0\\ 
wiwigapo & [VPOS 0.0-0.5] &left&0.1&0.0\\ 
fozumoba & [AREA 0.5-1.0]&large & 0.0&0.1\\ 
wefoto & [R 0.0-0.5]& low redness &0.2&0.2\\ 
togene & [R 0.5-1.0]& high redness &0.5&1.0\\ 
fumudanu & [G 0.0-0.5]& low greenness &0.2&0.2\\ 
puxedu & [G 0.5-1.0]& high greenness &0.4&0.4\\ 
\lspbottomrule
\end{tabular}
\caption{\label{tab:game50} First words after about a dozen games.}
\end{center}
\end{table}

@@

\begin{table}
\begin{center}
\begin{tabular}{ l  l  l  l  l }
\lsptoprule
{\itshape Word}&{\itshape Meaning}&{\itshape Translation} & {\bfshape a1}&{\bfshape a2} \\ \midrule
wawosido & [HPOS 0.0-0.5] &left&0.7&0.5\\ 
meluri & [HPOS 0.25-0.5] &medium left&0.4&0.4\\ 
gofubo & [HPOS 0.5-1.0]& right&1.0&1.0\\ 
vokomutu & [HPOS 0.5-0.75] &medium right&0.2&0.2\\ 
buwonipo & [HPOS 0.75-0.875] &strongly right&0.1&0.0\\ 
wiwigapo & [VPOS 0.0-0.5] &down&0.6&0.6\\ 
fozumoba & [AREA 0.5-1.0]&large & 0.2&0.4\\ 
wefoto & [R 0.0-0.5]& low redness &0.2&0.3\\ 
togene & [R 0.5-1.0]& high redness &1.0&1.0\\ 
fumudanu & [G 0.0-0.5]& low greenness &0.7&0.7\\ 
puxedu & [G 0.5-1.0]& high greenness &0.7&0.9\\ 
\lspbottomrule
\end{tabular}
\caption{\label{tab:goubo} Population lexicon after 100 games.}
\end{center}
\end{table}

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.75\textwidth]{chap7/figs/discri200}}
\caption{ \label{discri200} The discrimination trees 
of two embodied agents {\bfshape a1} (left) and {\bfshape a2} (right) 
after playing 100 games (top) and after 200 games (bottom).}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[width=0.8\columnwidth]{chap7/figs/plate-13}
\end{center}
\caption{ Top and bottom: Examples of new segmented scenes that stimulate conceptual development
and hence expansions of the lexicon. The most salient characteristic in the top scene is WIDTH and in the bottom 
scene HEIGHT.}
\label{fig:plate-13}
\end{figure}

@@

\begin{table}
\begin{center}
\begin{tabular}{ l  l  l }
\lsptoprule
{\itshape channel}& {\itshape obj-0} & {\itshape obj-1}\\ \midrule
HPOS & 0.08 & 0.09\\ 
VPOS & 0.21 & 0.10\\ 
HEIGHT & 0.10 & 0.15\\ 
WIDTH & 0.42 & 0.11\\ 
AREA & 0.34 & 0.16\\ 
R & 0.33 & 0.31\\ 
G & 0.68 & 0.65\\ 
B & 0.52 & 0.50\\ 
\lspbottomrule
\end{tabular}
\caption{ \label{tab:game126} Sensory data for game 126.}
\end{center}
\end{table}
Clearly width is the most salient channel for the speaker. 
It is therefore chosen and because the 
agents have already grown distinctions on this channel, 
discrimination succeeds and a new word can be
constructed and stored. 
\begin{verbatim}
Game 126 
  a2 is the speaker. a1 is the hearer. 
  a2 segments the context into 2 objects: 
       object-0 object-1
  a2 chooses object-0 as the topic 
  a2 categorises the topic as [WIDTH 0.5-1.0]
  a2 creates a new word: vaviwumu
  a2 says: vaviwumu
  a1 does not know vaviwumu
  a1 says: vaviwumu?
  a2 points to object-0
  a1 categorises the topic as [WIDTH 0.5-1.0]
  a1 stores vaviwumu as [WIDTH 0.5-1.0]
\end{verbatim}
\figref{fig:plate-13} (bottom) contains another example
of segmented images which have stimulated conceptual
growth and hence an expansion of the lexicon. 
In this case, categorisations on the HEIGHT channel are
relevant and a word developed for tall.

\begin{figure}
\begin{center}
\includegraphics[width=0.8\columnwidth]{chap7/figs/plate-14-area}
\end{center}
\caption{ Segmented images where distinctions on the AREA channel have been used. The topic (with dashed bounding 
box) in the two top cases has been categorised as large. In the bottom case, a word meaning `small' was used.}
\label{fig:plate-14}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[width=0.8\columnwidth]{chap7/figs/plate-15-vpos}
\end{center}
\caption{ The image segments from the top game require 
distinctions on the VPOS channel and 
caused the creation of a word for `upper'. The image 
segments in the middle lead to the use of refined
distinctions on the HPOS channel, so as to 
identify the middle square. The topic in the 
image segments at the bottom was done with a 
conjunction of categories: blue and tall.}
\label{fig:plate-15}
\end{figure}

@@

\begin{table}
\begin{center}
\begin{tabular}{ l  l  l  l  l }
\lsptoprule
{\itshape Word}&{\itshape Meaning}&{\itshape Translation} & {\bfshape a1}&{\bfshape a2} \\ \midrule
wawosido & [HPOS 0.0-0.5] &left&0.9&0.4\\ 
gixepo & [HPOS 0.0-0.25] & very left&0.4&0.4\\ 
wonuxa & [HPOS 0.0-0.25] & very left&0.10&0.0\\ 
meluri & [HPOS 0.25-0.5] &medium left&0.4&0.4\\ 
gofubo & [HPOS 0.5-1.0]& right&1.0&1.0\\ 
vokomutu & [HPOS 0.5-0.75] &medium right&0.4&0.4\\ 
buwonipo & [HPOS 0.75-0.875] &strongly right&0.1&0.0\\ 
wiwigapo & [VPOS 0.0-0.5] &down&0.5&0.4\\ 
putuwenu & [VPOS 0.5-1.0]&up & 0.5&0.5\\ 
vaviwumu & [WIDTH 0.5-1.0]&wide & 0.5&0.5\\ 
pesidumu & [AREA 0.0-0.5]&small& 0.2&0.2\\ 
fozumoba & [AREA 0.5-1.0]&large & 1.0&1.0\\ 
wefoto & [R 0.0-0.5]& low redness &0.2&0.1\\ 
togene & [R 0.5-1.0]& high redness &0.9&1.0\\ 
fumudanu & [G 0.0-0.5]& low greenness &0.7&0.9\\ 
puxedu & [G 0.5-1.0]& high greenness &0.9&0.9\\ 
\lspbottomrule
\end{tabular}
\caption{ \label{tab:upper} Population lexicon after 200 games.}
\end{center}
\end{table}
Note that new words have come into the lexicon 
for up (`putuwenu'), down (`wiwigapo'), and 
wide (`vaviwumu'). The discrimination trees of 
the WIDTH and HEIGHT channel have started to 
expand (\figref{discri200} bottom).

@@

Grounding language games in physical environments\is{perceptual anomalies}
makes it obviously harder
for the agents in a number of respects. First of 
all because perception and segmentation may differ, 
the salient characteristics of objects may not be the same
and as a result the hearer may guess another meaning for an 
unknown word, compared to 
the one used by the speaker. This happens for
example in the following game (game 128) which is based on 
the segmented images shown in \figref{fig:plate-16} (top). The
speaker's perception is shown to the left and the 
hearer's to the right. 
\begin{verbatim}
Game 128
  a1 is the speaker. a2 is the hearer. 
  a1 segments the context into 3 objects: 
       object-0 object-1 object-2
  a1 chooses object-0 as the topic 
  a1 categorises the topic as 
       [HPOS 0.0-0.25] [HPOS 0.0-0.125]
  a1 creates a new word: `bagaxe' for [HPOS 0.0-0.25]
  a1 says: bagaxe
  a2 does not know bagaxe
  a2 says: bagaxe?
  a1 points to object-0
  a2 categorises the topic as [AREA 0.5-1.0]
  a2 stores bagaxe as [AREA 0.5-1.0]
\end{verbatim}

\begin{figure}
\begin{center}
\includegraphics[width=0.8\columnwidth]{chap7/figs/plate-16}
\end{center}
\caption{ Examples of scenes causing confusion
due to perceptual anomalies. In the top scene, 
the segmentation of the speaker (left) makes the 
topic's most salient characteristic the horizontal
position, whereas for the hearer (right) the most salient
characteristic of the same segment is its area. In the 
bottom scene, speaker and hearer have segmented the 
scene into two, respectively three segments. This 
causes the game to fail even though both agents 
had the same meaning for the word used by the speaker.}
\label{fig:plate-16}
\end{figure}

@@

\begin{table}
\begin{center}
\begin{tabular}{ l  l  l  l }
\lsptoprule
{\itshape channel}& {\itshape obj-0} & {\itshape obj-1} & {\itshape obj-2}\\ 
HPOS & 0.03 & 0.15 & 0.33\\ 
VPOS & 0.29 & 0.29 & 0.01\\ 
HEIGHT & 0.37 & 0.39 & 0.07\\ 
WIDTH & 0.10 & 0.08 & 0.27\\ 
AREA & 0.29 & 0.22 & 0.16\\ 
R & 0.95 & 0.97 & 0.97 \\ 
G & 0.33 & 0.40 & 0.93\\ 
B & 0.35 & 0.44 & 0.26\\ 
\lspbottomrule
\end{tabular}
\caption{ \label{tab:game128} Sensory data for game 128.}
\end{center}
\end{table}
Context-scaling further
amplifies this difference which shows that it is not 
always beneficial to do so. {\bfshape a1} uses
the horizontal axis, creating a new word `bagaxe' 
meaning `very left' [HPOS 0.0-0.25] whereas {\bfshape a2}, for
whom the area is the most salient characteristic, 
associates this word with [AREA 0.5-1.0] (large). 
This example shows that grounding introduces 
additional risks for the introduction of semantic
incoherence in a group's lexicon.

Due to perceptual anomalies, a game may fail even though 
both agents already have the same meaning for the 
same word. This happens when this meaning yields different 
objects (or no objects at all) for the speaker and the hearer. 
As a consequence, the hearer adopts another meaning for 
the word which starts to compete with the one he already had. 
This happened in the following game (game 127)
based on the segmented images shown in \figref{fig:plate-16} (bottom). The speaker (left image) has identified
only two objects, but the hearer three. The 
speaker (left image) uses green as distinguishing 
characteristic which is indeed appropriate. but 
because the hearer's third object (the top rectangle) is 
also green, this distinction fails for him.  
Even though the hearer had already a well established
meaning for `puxedu', he associates a new meaning 
based on the WIDTH-channel with this word because this 
channel is now the most salient. 
\begin{verbatim}
Game 127
  a1 is the speaker. a2 is the hearer. 
  a1 segments the context into 2 objects: 
       s-object-0 s-object-1
  a1 chooses object-0 as the topic 
  a1 considers as salient G R AREA HPOS 
  a1 categorises the topic as 
       [HPOS 0.0-0.5] [HPOS 0.0-0.125]
       [AREA 0.0-0.5] [R 0.0-0.5] [G 0.5-1.0]
  a1 has the words
       puxedu for [G 0.5-1.0] (1.0)
       wawosido for [HPOS 0.0-0.5] (0.7)
       pesidimu for [AREA 0.0-0.5] (0.2)
       wefoto for [R 0.0-0.5] (0.2)
       buwonipo for [G 0.5-1.0] (0.0)
  a1 says: puxedu
  a2 segments the context into 3 objects: 
       h-object-0 h-object-1 h-object-2
  a2 interprets puxedu as
       [G 0.5-1.0] (0.20)
  a2 identifies h-object-0 h-object-1
  a2 says: puxedu?
  a1 points to s-object-0
  a2 categorises the topic as [WIDTH 0.5-1.0]
  a2 stores puxedu as [WIDTH 0.5-1.0]
\end{verbatim}

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.65\textwidth]{chap7/figs/landscape}}
\caption{ \label{RMF1} Typical segment of a semiotic
landscape capturing the co-occurrence relations between
referents, meanings and word forms. The FM/MF relations are
in thick lines, the RM/MR relations in regular lines, and 
the RF/FR relations in dashed lines.}
\end{figure}
The relations in the landscape are labeled as in 
the following table:
\begin{center}
\begin{tabular}{ l  l  }
\lsptoprule
{\itshape Label}& {\itshape Relation } \\ \midrule
RM & referent to meaning \\ 
MR & meaning to referent \\ 
FM & form to meaning\\ 
MF & meaning to form\\ 
RF & referent to form \\ 
FR & form to referent  \\ 
\lspbottomrule
\end{tabular}
\end{center}

>>>>

%\documentclass{../_combined/fcg-book}
\chapter{Grounding} \label{chap:7}

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.50\textwidth]{chap7/figs/sieve3.pdf}}
\caption{\label{sieve3}Flow of solutions through 
the different cognitive layers when an utterance is being produced.
Re-entry is necessary because the decision at each layer depends on
decisions at subsequent layers.} 
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[width=0.8\columnwidth]{chap7/figs/plate-10.pdf}
\end{center}
\caption{Three examples of segmented images. The 
topic is indicated by a dashed bounding box in the 
image of the speaker. Segments which are too small 
are ignored. The topics have all been conceptualised
as being `to the right' and so the same word 
`gofubo' has been used to refer to them. }
\label{fig:plate-10}
\end{figure}

@@

\begin{table}
\begin{center}
\begin{tabular}{ l  l  l }
\lsptoprule
{\itshape channel}& {\itshape obj-0} & {\itshape obj-1}\\ \midrule
HPOS & 0.27 & 0.16\\ 
VPOS & 0.20 & 0.20\\ 
HEIGHT & 0.15 & 0.15\\ 
WIDTH & 0.10 & 0.11\\ 
AREA & 0.10 & 0.10\\ 
R & 0.23 & 0.25\\ 
G & 0.32 & 0.34\\ 
B & 0.63 & 0.65\\ 
\lspbottomrule
\end{tabular}
\caption{\label{tab:game3b}First words after about a dozen games.}
\end{center}
\end{table}

HPOS is the most salient channel. After sensor-scaling, 
the two values for HPOS are still be drawn further
apart with 1.0 for object-0 and 
0.0 for object-1 so that the category [HPOS 0.5–1.0] easily 
distinguishes the topic (object-0) from object-1.

The left image is that of the hearer {\bfshape a1}, the right one
that of the speaker, {\bfshape a2}.
{\bfshape a2} had already invented the word `gofubo' for 
[HPOS 0.5–1.0] (to the right) in an earlier game, 
but the word was not acquired in that game
by {\bfshape a1} because he still missed the 
appropriate distinction. Meanwhile the HPOS category 
is available to {\bfshape a1} due to an
expansion of his discrimination networks
and so he can store the word: 
\begin{verbatim}
Game 3 
  a2 is the speaker. a1 is the hearer. 
  a2 segments the context into 2 objects: 
       object-0 object-1
  a2 chooses object-0 as the topic 
  a2 categorises the topic as [HPOS 0.5–1.0]
  a2 says: gofubo
  a1 does not know gofubo
  a1 says: gofubo?
  a2 points to object-0
  a1 categorises the topic as [HPOS 0.5–1.0]
  a1 stores gofubo as [HPOS 0.5–1.0]
\end{verbatim}
This game proceeds essentially like in the computer
simulations studied before. The major difference is
that now real images have been used, the objects
considered are the outcome of segmentation processes,
the sensory characteristics have been derived from the 
image itself, and the pointing has been done by 
physically moving the cameras.

@@

Immediately the agents apply this word to very 
different scenes, such as the ones in \figref{fig:plate-10} (middle and bottom). The middle picture
shows on the left the segmented scene
from the speaker in game 5, and on the right the one
from the hearer. In game 5, the figures are blue
rectangles. They are much further apart than 
the two circles in game 3. Nevertheless, 
after scaling, they are categorised and
conceptualised similarly and therefore
the same word could be used effectively. The concept 
of [HPOS 0.5–1.0] and hence the word `gofubo' 
is general {\itshape from the very start}. 
The agents do not need to see many examples because they do 
not use inductive generalisation. Instead, 
the HPOS category is constructed in a top-down fashion 
as soon as the HPOS channel has been salient
and is immediately available for use in the 
discrimination game and hence in verbalising 
the scene. This explains why the word-meaning acquisition
process observed in the experiments goes so 
amazingly fast.

The bottom of \figref{fig:plate-10}
shows yet another scene where the word `gofubo' 
was used with success. The topic is the rightmost
shape in the scene and so [HPOS 0.5–1.0] is once more 
distinctive. To an outside observer it may look 
like the agents have performed a gigantic inductive
leap, but this is not the case at all. The agents
do not try to abstract the commonalities from 
different examples but construct distinctions in a top 
down fashion and try to apply them to the perceived image.

@@

\begin{table}
\begin{center}
\begin{tabular}{ l  l  l  l  l }
\lsptoprule
{\itshape Word}&{\itshape Meaning}&{\itshape Translation} & {\bfshape a1}&{\bfshape a2} \\ \midrule
wawosido & [HPOS 0.0–0.5] &left&0.4&0.4\\ 
meluri & [HPOS 0.25–0.5] &medium left&0.1&0.0\\ 
gofubo & [HPOS 0.5–1.0]& right&1.0&1.0\\ 
wiwigapo & [VPOS 0.0–0.5] &left&0.1&0.0\\ 
fozumoba & [AREA 0.5–1.0]&large & 0.0&0.1\\ 
wefoto & [R 0.0–0.5]& low redness &0.2&0.2\\ 
togene & [R 0.5–1.0]& high redness &0.5&1.0\\ 
fumudanu & [G 0.0–0.5]& low greenness &0.2&0.2\\ 
puxedu & [G 0.5–1.0]& high greenness &0.4&0.4\\ 
\lspbottomrule
\end{tabular}
\caption{\label{tab:game50}First words after about a dozen games.}
\end{center}
\end{table}

@@

\begin{table}
\begin{center}
\begin{tabular}{ l  l  l  l  l }
\lsptoprule
{\itshape Word}&{\itshape Meaning}&{\itshape Translation} & {\bfshape a1}&{\bfshape a2} \\ \midrule
wawosido & [HPOS 0.0–0.5] &left&0.7&0.5\\ 
meluri & [HPOS 0.25–0.5] &medium left&0.4&0.4\\ 
gofubo & [HPOS 0.5–1.0]& right&1.0&1.0\\ 
vokomutu & [HPOS 0.5–0.75] &medium right&0.2&0.2\\ 
buwonipo & [HPOS 0.75–0.875] &strongly right&0.1&0.0\\ 
wiwigapo & [VPOS 0.0–0.5] &down&0.6&0.6\\ 
fozumoba & [AREA 0.5–1.0]&large & 0.2&0.4\\ 
wefoto & [R 0.0–0.5]& low redness &0.2&0.3\\ 
togene & [R 0.5–1.0]& high redness &1.0&1.0\\ 
fumudanu & [G 0.0–0.5]& low greenness &0.7&0.7\\ 
puxedu & [G 0.5–1.0]& high greenness &0.7&0.9\\ 
\lspbottomrule
\end{tabular}
\caption{\label{tab:goubo}Population lexicon after 100 games.}
\end{center}
\end{table}

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.75\textwidth]{chap7/figs/discri200.pdf}}
\caption{\label{discri200}The discrimination trees 
of two embodied agents {\bfshape a1} (left) and {\bfshape a2} (right) 
after playing 100 games (top) and after 200 games (bottom).}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[width=0.8\columnwidth]{chap7/figs/plate-13.pdf}
\end{center}
\caption{Top and bottom: Examples of new segmented scenes that stimulate conceptual development
and hence expansions of the lexicon. The most salient characteristic in the top scene is WIDTH and in the bottom 
scene HEIGHT.}
\label{fig:plate-13}
\end{figure}

@@

\begin{table}
\begin{center}
\begin{tabular}{ l  l  l }
\lsptoprule
{\itshape channel}& {\itshape obj-0} & {\itshape obj-1}\\ \midrule
HPOS & 0.08 & 0.09\\ 
VPOS & 0.21 & 0.10\\ 
HEIGHT & 0.10 & 0.15\\ 
WIDTH & 0.42 & 0.11\\ 
AREA & 0.34 & 0.16\\ 
R & 0.33 & 0.31\\ 
G & 0.68 & 0.65\\ 
B & 0.52 & 0.50\\ 
\lspbottomrule
\end{tabular}
\caption{\label{tab:game126}Sensory data for game 126.}
\end{center}
\end{table}
Clearly width is the most salient channel for the speaker. 
It is therefore chosen and because the 
agents have already grown distinctions on this channel, 
discrimination succeeds and a new word can be
constructed and stored. 
\begin{verbatim}
Game 126 
  a2 is the speaker. a1 is the hearer. 
  a2 segments the context into 2 objects: 
       object-0 object-1
  a2 chooses object-0 as the topic 
  a2 categorises the topic as [WIDTH 0.5–1.0]
  a2 creates a new word: vaviwumu
  a2 says: vaviwumu
  a1 does not know vaviwumu
  a1 says: vaviwumu?
  a2 points to object-0
  a1 categorises the topic as [WIDTH 0.5–1.0]
  a1 stores vaviwumu as [WIDTH 0.5–1.0]
\end{verbatim}
\figref{fig:plate-13} (bottom) contains another example
of segmented images which have stimulated conceptual
growth and hence an expansion of the lexicon. 
In this case, categorisations on the HEIGHT channel are
relevant and a word developed for tall.

\begin{figure}
\begin{center}
\includegraphics[width=0.8\columnwidth]{chap7/figs/plate-14-area.pdf}
\end{center}
\caption{Segmented images where distinctions on the AREA channel have been used. The topic (with dashed bounding 
box) in the two top cases has been categorised as large. In the bottom case, a word meaning `small' was used.}
\label{fig:plate-14}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[width=0.8\columnwidth]{chap7/figs/plate-15-vpos.pdf}
\end{center}
\caption{The image segments from the top game require 
distinctions on the VPOS channel and 
caused the creation of a word for `upper'. The image 
segments in the middle lead to the use of refined
distinctions on the HPOS channel, so as to 
identify the middle square. The topic in the 
image segments at the bottom was done with a 
conjunction of categories: blue and tall.}
\label{fig:plate-15}
\end{figure}

@@

\begin{table}
\begin{center}
\begin{tabular}{ l  l  l  l  l }
\lsptoprule
{\itshape Word}&{\itshape Meaning}&{\itshape Translation} & {\bfshape a1}&{\bfshape a2} \\ \midrule
wawosido & [HPOS 0.0–0.5] &left&0.9&0.4\\ 
gixepo & [HPOS 0.0–0.25] & very left&0.4&0.4\\ 
wonuxa & [HPOS 0.0–0.25] & very left&0.10&0.0\\ 
meluri & [HPOS 0.25–0.5] &medium left&0.4&0.4\\ 
gofubo & [HPOS 0.5–1.0]& right&1.0&1.0\\ 
vokomutu & [HPOS 0.5–0.75] &medium right&0.4&0.4\\ 
buwonipo & [HPOS 0.75–0.875] &strongly right&0.1&0.0\\ 
wiwigapo & [VPOS 0.0–0.5] &down&0.5&0.4\\ 
putuwenu & [VPOS 0.5–1.0]&up & 0.5&0.5\\ 
vaviwumu & [WIDTH 0.5–1.0]&wide & 0.5&0.5\\ 
pesidumu & [AREA 0.0–0.5]&small& 0.2&0.2\\ 
fozumoba & [AREA 0.5–1.0]&large & 1.0&1.0\\ 
wefoto & [R 0.0–0.5]& low redness &0.2&0.1\\ 
togene & [R 0.5–1.0]& high redness &0.9&1.0\\ 
fumudanu & [G 0.0–0.5]& low greenness &0.7&0.9\\ 
puxedu & [G 0.5–1.0]& high greenness &0.9&0.9\\ 
\lspbottomrule
\end{tabular}
\caption{\label{tab:upper}Population lexicon after 200 games.}
\end{center}
\end{table}
Note that new words have come into the lexicon 
for up (`putuwenu'), down (`wiwigapo'), and 
wide (`vaviwumu'). The discrimination trees of 
the WIDTH and HEIGHT channel have started to 
expand (\figref{discri200} bottom).

@@

Grounding language games in physical environments\is{perceptual anomalies}
makes it obviously harder
for the agents in a number of respects. First of 
all because perception and segmentation may differ, 
the salient characteristics of objects may not be the same
and as a result the hearer may guess another meaning for an 
unknown word, compared to 
the one used by the speaker. This happens for
example in the following game (game 128) which is based on 
the segmented images shown in \figref{fig:plate-16} (top). The
speaker's perception is shown to the left and the 
hearer's to the right. 
\begin{verbatim}
Game 128
  a1 is the speaker. a2 is the hearer. 
  a1 segments the context into 3 objects: 
       object-0 object-1 object-2
  a1 chooses object-0 as the topic 
  a1 categorises the topic as 
       [HPOS 0.0–0.25] [HPOS 0.0–0.125]
  a1 creates a new word: `bagaxe' for [HPOS 0.0–0.25]
  a1 says: bagaxe
  a2 does not know bagaxe
  a2 says: bagaxe?
  a1 points to object-0
  a2 categorises the topic as [AREA 0.5–1.0]
  a2 stores bagaxe as [AREA 0.5–1.0]
\end{verbatim}

\begin{figure}
\begin{center}
\includegraphics[width=0.8\columnwidth]{chap7/figs/plate-16.pdf}
\end{center}
\caption{Examples of scenes causing confusion
due to perceptual anomalies. In the top scene, 
the segmentation of the speaker (left) makes the 
topic's most salient characteristic the horizontal
position, whereas for the hearer (right) the most salient
characteristic of the same segment is its area. In the 
bottom scene, speaker and hearer have segmented the 
scene into two, respectively three segments. This 
causes the game to fail even though both agents 
had the same meaning for the word used by the speaker.}
\label{fig:plate-16}
\end{figure}

@@

\begin{table}
\begin{center}
\begin{tabular}{ l  l  l  l }
\lsptoprule
{\itshape channel}& {\itshape obj-0} & {\itshape obj-1} & {\itshape obj-2}\\ 
HPOS & 0.03 & 0.15 & 0.33\\ 
VPOS & 0.29 & 0.29 & 0.01\\ 
HEIGHT & 0.37 & 0.39 & 0.07\\ 
WIDTH & 0.10 & 0.08 & 0.27\\ 
AREA & 0.29 & 0.22 & 0.16\\ 
R & 0.95 & 0.97 & 0.97 \\ 
G & 0.33 & 0.40 & 0.93\\ 
B & 0.35 & 0.44 & 0.26\\ 
\lspbottomrule
\end{tabular}
\caption{\label{tab:game128}Sensory data for game 128.}
\end{center}
\end{table}
Context-scaling further
amplifies this difference which shows that it is not 
always beneficial to do so. {\bfshape a1} uses
the horizontal axis, creating a new word `bagaxe' 
meaning `very left' [HPOS 0.0–0.25] whereas {\bfshape a2}, for
whom the area is the most salient characteristic, 
associates this word with [AREA 0.5–1.0] (large). 
This example shows that grounding introduces 
additional risks for the introduction of semantic
incoherence in a group's lexicon.

Due to perceptual anomalies, a game may fail even though 
both agents already have the same meaning for the 
same word. This happens when this meaning yields different 
objects (or no objects at all) for the speaker and the hearer. 
As a consequence, the hearer adopts another meaning for 
the word which starts to compete with the one he already had. 
This happened in the following game (game 127)
based on the segmented images shown in \figref{fig:plate-16} (bottom). The speaker (left image) has identified
only two objects, but the hearer three. The 
speaker (left image) uses green as distinguishing 
characteristic which is indeed appropriate. but 
because the hearer's third object (the top rectangle) is 
also green, this distinction fails for him.  
Even though the hearer had already a well established
meaning for `puxedu', he associates a new meaning 
based on the WIDTH-channel with this word because this 
channel is now the most salient. 
\begin{verbatim}
Game 127
  a1 is the speaker. a2 is the hearer. 
  a1 segments the context into 2 objects: 
       s-object-0 s-object-1
  a1 chooses object-0 as the topic 
  a1 considers as salient G R AREA HPOS 
  a1 categorises the topic as 
       [HPOS 0.0–0.5] [HPOS 0.0–0.125]
       [AREA 0.0–0.5] [R 0.0–0.5] [G 0.5–1.0]
  a1 has the words
       puxedu for [G 0.5–1.0] (1.0)
       wawosido for [HPOS 0.0–0.5] (0.7)
       pesidimu for [AREA 0.0–0.5] (0.2)
       wefoto for [R 0.0–0.5] (0.2)
       buwonipo for [G 0.5–1.0] (0.0)
  a1 says: puxedu
  a2 segments the context into 3 objects: 
       h-object-0 h-object-1 h-object-2
  a2 interprets puxedu as
       [G 0.5–1.0] (0.20)
  a2 identifies h-object-0 h-object-1
  a2 says: puxedu?
  a1 points to s-object-0
  a2 categorises the topic as [WIDTH 0.5–1.0]
  a2 stores puxedu as [WIDTH 0.5–1.0]
\end{verbatim}

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.65\textwidth]{chap7/figs/landscape.pdf}}
\caption{\label{RMF1}Typical segment of a semiotic
landscape capturing the co-occurrence relations between
referents, meanings and word forms. The FM/MF relations are
in thick lines, the RM/MR relations in regular lines, and 
the RF/FR relations in dashed lines.}
\end{figure}
The relations in the landscape are labeled as in 
the following table:
\begin{center}
\begin{tabular}{ l  l  }
\lsptoprule
{\itshape Label}& {\itshape Relation } \\ \midrule
RM & referent to meaning \\ 
MR & meaning to referent \\ 
FM & form to meaning\\ 
MF & meaning to form\\ 
RF & referent to form \\ 
FR & form to referent  \\ 
\lspbottomrule
\end{tabular}
\end{center}

The partial landscape in \figref{RMF1} (taken from
an actual experiment) contains an example where the agents use
two possible meanings for conceptualising {\itshape situation2}, namely
[GRAY 0.0–0.25] (very light) and [HPOS 0.5–0.75]
(medium right ). The words `katapu' and `tisame' lexicalise
[GRAY 0.0–0.25] and `wobo' and `tisame' [HPOS 0.5–0.75]. 
Each meaning has therefore two synonyms and `tisame' is 
ambiguous; it can mean both [GRAY 0.0–0.25]
and [HPOS 0.5–0.75]. Three words
are used to communicate the situation: `katapu', `tisame', 
and `wobo'. Such a structure is typical for grounded
lexicon evolution and complexity rapidly increases when the
same meanings are used to communicate about other
situations (which is obviously very common and indeed desirable).
